---
title: 'Status Report #4 - Romano'
output: pdf_document
date: "2025-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setup

```{r prepare}
# Load in libraries 
library(MASS)
library(e1071)
library(rpart)
library(randomForest)

# Load data and view it (make sure everything looks right)
HeartAttack <- read.csv("HeartAttack.csv")
head(HeartAttack)

# Load cleaned data from Status Report #1 and view it (make sure it looks right)
HeartAttackClean <- readRDS("HeartAttackClean.rds")
head(HeartAttackClean)
```
Logistic Regression Model - Decision Tree

```{r logistical_regression_dt}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")

# Split into training and testing sets 70/30
set.seed(1)
train_idx <- sample(1:nrow(HeartAttackClean), 0.7 * nrow(HeartAttackClean))
train <- HeartAttackClean[train_idx, ]
test  <- HeartAttackClean[-train_idx, ]

# Fit logistic regression
fit <- glm(
  reformulate(intersect(predictors, names(HeartAttackClean)), response = "Outcome"),
  data = HeartAttackClean,
  family = binomial(),
  na.action = na.omit
)

# Add decision tree enhancement 
tree_fit <- rpart(Outcome ~ ., data = train, method = "class"); tree_pred <- predict(tree_fit, newdata = test, type = "class")

# Results
summary(fit)          # coefficients (log-odds), p-values
exp(coef(fit))        # odds ratios

# Simple in-sample predictions & accuracy (threshold = 0.5)
p    <- predict(fit, type = "response")
pred <- as.integer(p >= 0.5)
table(truth = HeartAttackClean$Outcome, pred = pred)
mean(pred == HeartAttackClean$Outcome)  # accuracy

mean(tree_pred == test$Outcome)
```

Logistic Regression Model - Random Forest

```{r logistical_regression_rf}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")

# Split into training and testing sets (70/30 for example)
set.seed(1)
train_idx <- sample(1:nrow(HeartAttackClean), 0.7 * nrow(HeartAttackClean))
train <- HeartAttackClean[train_idx, ]
test  <- HeartAttackClean[-train_idx, ]

# Fit logistic regression
fit <- glm(
  reformulate(intersect(predictors, names(HeartAttackClean)), response = "Outcome"),
  data = HeartAttackClean,
  family = binomial(),
  na.action = na.omit
)

# Results
summary(fit)          # coefficients (log-odds), p-values
exp(coef(fit))        # odds ratios

# Simple in-sample predictions & accuracy (threshold = 0.5)
p    <- predict(fit, type = "response")
pred <- as.integer(p >= 0.5)
table(truth = HeartAttackClean$Outcome, pred = pred)
mean(pred == HeartAttackClean$Outcome)  # accuracy

# Add random forest enhancement 
#rf_fit <- randomForest(Outcome ~ ., data = na.omit(train), ntree = 500)
#rf_pred <- predict(rf_fit, newdata = na.omit(test))
## I need some help with this. It keeps buffering, and it will not load results 

```

LDA Model - Decsion Tree

```{r LDA_dt}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")


# Split into training and test sets
set.seed(1)
n <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8 * n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

# Fit LDA model
lda_fit <- lda(Outcome ~ Age + Gender + FamilyHistory + Ethnicity +
               Smoker + BloodPressure + Cholesterol + BMI + Diabetes +
               Diet + PhysicalActivity + HeartRate + AlcoholConsumption,
               data = train)

# Add decision tree enhancement 
tree_fit <- rpart(Outcome ~ ., data = train, method = "class"); tree_pred <- predict(tree_fit, newdata = test, type = "class")


# Predict on test data
lda_pred <- predict(lda_fit, newdata = test)$class

# Confusion matrix and accuracy
lda_cm <- table(truth = test$Outcome, pred = lda_pred)
lda_acc <- mean(lda_pred == test$Outcome)

lda_cm
lda_acc

mean(tree_pred == test$Outcome)
```

LDA Model - Random Forest

```{r LDA_rf}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")


# Split into training and test sets
set.seed(1)
n <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8 * n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

# Fit LDA model
lda_fit <- lda(Outcome ~ Age + Gender + FamilyHistory + Ethnicity +
               Smoker + BloodPressure + Cholesterol + BMI + Diabetes +
               Diet + PhysicalActivity + HeartRate + AlcoholConsumption,
               data = train)

# Add random forest enhancement 
#rf_fit <- randomForest(Outcome ~ ., data = train, ntree = 500, importance = TRUE); rf_pred <- predict(rf_fit, newdata = test)
## Same issue as above 

# Predict on test data
lda_pred <- predict(lda_fit, newdata = test)$class

# Confusion matrix and accuracy
lda_cm <- table(truth = test$Outcome, pred = lda_pred)
lda_acc <- mean(lda_pred == test$Outcome)

lda_cm
lda_acc
```

QDA Model - Decision Tree 

```{r QDA_dt}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")

set.seed(1)
n  <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8*n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

train  <- train[complete.cases(train[, c("Outcome", predictors)]), ]
test_cc <- test[complete.cases(test[, predictors]), ]

qda_fit  <- qda(as.formula(paste("Outcome ~", paste(predictors, collapse = " + "))), data = train)
qda_pred <- predict(qda_fit, newdata = test_cc)$class

# Add decsion tree enhancement 
tree_fit <- rpart(Outcome ~ ., data = train, method = "class"); tree_pred <- predict(tree_fit, newdata = test, type = "class")

qda_cm  <- table(truth = test_cc$Outcome, pred = qda_pred)
qda_acc <- mean(qda_pred == test_cc$Outcome)

qda_cm
qda_acc

mean(tree_pred == test$Outcome)
```

QDA Model - Random Forest 

```{r QDA_rf}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")

set.seed(1)
n  <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8*n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

train  <- train[complete.cases(train[, c("Outcome", predictors)]), ]
test_cc <- test[complete.cases(test[, predictors]), ]

qda_fit  <- qda(as.formula(paste("Outcome ~", paste(predictors, collapse = " + "))), data = train)
qda_pred <- predict(qda_fit, newdata = test_cc)$class

# Add random forest enhancement 

qda_cm  <- table(truth = test_cc$Outcome, pred = qda_pred)
qda_acc <- mean(qda_pred == test_cc$Outcome)

qda_cm
qda_acc
```

Naive Bayes Model - Decision Tree

```{r Naive_Bayes_dt}
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")
set.seed(1)
n  <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8*n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

train  <- train[complete.cases(train[, c("Outcome", predictors)]), ]
test_cc <- test[complete.cases(test[, predictors]), ]

nb_fit  <- naiveBayes(as.formula(paste("Outcome ~", paste(predictors, collapse = " + "))), data = train)
nb_pred <- predict(nb_fit, newdata = test_cc, type = "class")

# Add decision tree enhancement 
tree_fit <- rpart(Outcome ~ ., data = train, method = "class"); tree_pred <- predict(tree_fit, newdata = test, type = "class")

nb_cm  <- table(truth = test_cc$Outcome, pred = nb_pred)
nb_acc <- mean(nb_pred == test_cc$Outcome)

nb_cm
nb_acc

mean(tree_pred == test$Outcome)
```

Naive Bayes Model - Random Forest 

```{r Naive_Bayes_rf}
test$Outcome <- as.factor(test$Outcome)
# Choose predictors 
predictors <- c("Age","Gender","FamilyHistory","Ethnicity",
                "Smoker","BloodPressure","Cholesterol","BMI",
                "Diabetes","Diet","PhysicalActivity","HeartRate",
                "AlcoholConsumption")
set.seed(1)
n  <- nrow(HeartAttackClean)
id <- sample(n, floor(0.8*n))
train <- HeartAttackClean[id, ]
test  <- HeartAttackClean[-id, ]

train  <- train[complete.cases(train[, c("Outcome", predictors)]), ]
test_cc <- test[complete.cases(test[, predictors]), ]

nb_fit  <- naiveBayes(as.formula(paste("Outcome ~", paste(predictors, collapse = " + "))), data = train)
nb_pred <- predict(nb_fit, newdata = test_cc, type = "class")

# Add random forest enhancement 

nb_cm  <- table(truth = test_cc$Outcome, pred = nb_pred)
nb_acc <- mean(nb_pred == test_cc$Outcome)

nb_cm
nb_acc
```

Conclusion:  
- Adding decsion trees to my models did not improve the accuracy as I was hoping it would  
- I was looking for a durastic difference in accuracy percentage, but there was little to noo change  
- I am hopeful that adding a random forest enhancement will improve my accuracy, but I was not able to get that to work with my data after a lot of research