---
title: 'Status Report #2'
output:
  html_document:
    df_print: paged
date: "2025-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tasks: Create EDR, correlations, initial visualizations

```{r prepare}
# Load necessary libraries
library(readr)
library(DiagrammeR)
library(dplyr)

# Load data and view it (make sure everything looks right)
HeartAttack <- read.csv("HeartAttack.csv")
head(HeartAttack)

# Load cleaned data from Status Report #1 and view it (make sure it looks right)
HeartAttackClean <- readRDS("HeartAttackClean.rds")
head(HeartAttackClean)
```

```{r ERD}
# Create ERD
mermaid("
erDiagram
  HeartAttack {
    int Age
    int Gender
    int Cholesterol
    int BloodPressure
    int HeartRate
    int BMI
    int Smoker
    int Diabetes
    int Hypertension
    int FamilyHistory
    int PhysicalActivity
    int AlcoholConsumption
    int Diet
    int StressLevel
    int Ethnicity
    int Income
    int EducationLevel
    int Medication
    int ChectPainType
    int ECGResults
    int MaxHeartRate
    int ST_Depression
    int ExerciseInducedAngina
    int Slope
    int NumberOfMajorVessels
    int Thalassemia
    int PreviousHeartAttack
    int StrokeHistory
    int Residence
    int EmploymentStatus
    int MaritalStatus
    int Outcome
  }
")


```

```{r chunk_correlations}
# These are the variables I want to see a correlation with Outcome
vars <- c("Age","Gender","FamilyHistory","Smoker",
          "PhysicalActivity","AlcoholConsumption",
          "Diet","StressLevel","EducationLevel")

# I want to see if the above variables have correlations with Outcome 
target <- "Outcome"   # <-- change if your column is named differently

df <- HeartAttackClean %>% select(any_of(c(vars, target)))

# Compute correlation with Outcome
res <- lapply(setdiff(names(df), target), function(v) {
  x <- df[[v]]
  y <- df[[target]]
  ok <- is.finite(df[[v]]) & is.finite(y)
  r  <- suppressWarnings(cor(df[[v]][ok], y[ok], method = "pearson"))  # phi for 0/1
  data.frame(variable = v,
             r = r,
             direction = ifelse(is.na(r), NA,
                                ifelse(r > 0, "positive",
                                       ifelse(r < 0, "negative", "zero"))))
})
res <- bind_rows(res) %>% arrange(match(variable, vars))
print(res)
```
- Positive: More heart attacks than no heart attacks  
- Negative: Fewer heart attacks than heart attacks  
- Age and stress level have less heart attack outcomes  
- Gender, family history, smoker, physical activity, alcohol consumption, diet, and education level have more heart attack outcomes  
- This tells us that the positive factors can increase the risk of a heart attack 

```{r, fig.width=9, fig.height=6}
# Create initial plots
barplot(table(HeartAttackClean$Smoker), main="Smoker", xlab="0/1", ylab="Count", col = c("pink", "lightblue"))     # Smoker vs Non-Smoker Count
barplot(table(HeartAttackClean$FamilyHistory), main="FamilyHistory", xlab="0/1", ylab="Count", col = c("purple", "steelblue"))     # Family History Yes vs No Count
barplot(table(HeartAttackClean$Gender), main="Gender", xlab="0/1", ylab="Count", col = c("pink", "blue"))          # Male vs Female Count
barplot(table(HeartAttackClean$PhysicalActivity), main="PhysicalActivity", xlab="level", ylab="Count", col = c("aquamarine", "cadetblue2", "coral1", "darkolivegreen", "deeppink", "orchid", "royalblue1"))        # Physical Activity Count

plot(HeartAttackClean$Age, HeartAttackClean$Cholesterol,
     xlab="Age", ylab="Cholesterol", pch=16, main="Age vs Cholesterol", col = c("skyblue", "slateblue1"))       # Age vs Cholesterol
plot(HeartAttackClean$BMI, HeartAttackClean$BloodPressure,
     xlab="BMI", ylab="BloodPressure", pch=16, main="BMI vs BloodPressure", col = c("thistle1", "violetred1"))      # BMI vs Blood Pressure 

```

